---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pcoskbR

pcoskbR provides an R interface to acess datasets related to PCOS in PCOSKBR2. As well as, data mining tools for comorbidity prediction and identification of enriched pathways and hub genes.

## Installation

You can install the released version of pcoskbR from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("pcoskbR")
```
## Getting pages and filters you need to search for  
  Here we show the list of pages in PCOSKBR2 we can search for:
```{r}
library(pcoskbR)
listPages()
```
  Let's say we want to search among the gene and we want to know the filters for Genes
  Here we show the list of pages in PCOSKBR2 we can search for. 
```{r}
listFilters(page = "Genes")
```
## Searching for information about PCOS
  Here we select the page and the filter to get a dataframe of the result from PCOSKBR2.
```{r}

manually_curated_genes = browse(page = "Genes", filter = "Manually curated")
head(manually_curated_genes)
```


